---
title: "power_avocado"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(data.table)
library(stargazer)
library(dplyr)
library(tidyverse)
library(patchwork)
library(sandwich)
library(reshape2)
library(car)
library(lmtest)#coeftest()
library(bcp)
library(changepoint)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
getChangeHUE <- function(df) {
  #this function is called after 
  d_temp <- as.data.frame(df)
  #this is the perent of hue information is stored
  dg <- d_temp$value
  set.seed(1000)
  bcp_x <- bcp(dg, return.mcmc = TRUE)
  #plot(bcp_x)
  
  bcp_sum <- as.data.frame(summary(bcp_x))
  bcp_sum$id <- 1:length(dg)
  #selecting draft_number with posterior probabltiy greater than 0.2
  sel <- bcp_sum[which(bcp_x$posterior.prob > 0.2), ]
  #by this time all the bananas went bad
  sel <- sel[sel$id <49,]
  sel <- sel[which.max(sel$Probability),]
  #get the id
  time_of_change <- time(dg)[sel$id]
  return(time_of_change[1])
}

getIncre <- function(df) {
  #takes avocado cumuluative hue value
  #returns frequency for each value of hue
  #which will be used to create sample data
  max <- max(df)
  min <- min(df)
  diff <- max-min
  
  hue_index <- 0:49
  len <- dim(df)[1]
  H_vector <- vector()
  
  for(i in 1:len){
     #set.seed(i)
     if (i == 1){
        #initial value is 0, the reference point
        H_vector[1] <- 0
     }
     else{
     #H_vector[i] <- (df[i]-df[i-1])*100
     H_vector[i] <- (df[i]-df[i-1])
     }
  }
  result <- H_vector
  #result1 <- cbind(hue_index,H_vector)
  #result <- get_ind_data(result1)
  return(result)
}

#create sample data based on pdf
get_ind_data <- function(df) {
 sample <- vector()
 len <- dim(df)[1]
 for(i in 1:len){
    add_it <- rep(df[i,1][[1]],round(df[i,2][[1]],2)*100)
    sample <- c(sample,add_it)
 }
return(sample)
}


#################################################
get_ks_permutation <- function(control,treat,sim) {
  #control is sample data for control
  #treat is the sampe data for treat,
  control <- control
  treatment <- treat
  
  Z <- c(control,treatment)
  n <- length(control )
  m <- length(treatment)
  N <- length(Z)
  
  # Number of permutations
  K = sim

  # Test statistic
  getD <- function(A,B) ks.test(A,B)$statistic
  
  # Test statistic for the observed sample
  maxD <- getD(control,treatment)
  maxD
  # Vector of test statistics for each permutation
  TT <- vector()
  
  # Permutation test
  for(i in 1:K){
    set.seed(i)
    Z.pi <- sample(Z, N, replace = FALSE)
    TT[i] <- getD(Z.pi[1:n], Z.pi[(n+1):(n+m)])
  }
  
  # Visualising the permuted test statistics
  hist(TT,xlim = c(0,0.5))
  abline(v = maxD, lwd = 2, col = "green")
  box()
  
  mean(TT>maxD)
}

```


# AVOCADO

## Block and randomization experimental design. 

- Three blocks and 12 subjects each

```{r warning=FALSE}
## Read and format data
da <- fread('avocado_blackness.csv')
#convert wide to long extract data attribute 
#and create new columns
df <-(melt(da, id.vars=c("hue_index")))
df$block<- substring(df$variable,0,1)
df$treatment <- substring(df$variable,2,2)
df$id <- substring(df$variable,3,4)
d_avo_raw <- df[,c("block","treatment","id","hue_index","value")]
```

- Figure A1 shows empirical cumulative distribution of black ratio as function of hue
- Figure A1 shows suggests that there exist both fixed effect for block and individual level.
  - Having differnt background when taking picture of avocado or differences in 
  lighting could have caused the fixed effect. 

```{r, warning=FALSE, fig.cap="Cumulative distribution of black ratio"}
d_avo_raw %>% ggplot(aes(x = hue_index, y = value, color=treatment)) + 
  geom_point(aes(color = treatment)) + facet_wrap(~block,ncol = 1) + 
  xlim(20,35) + 
  xlab("Hue") + ylab("Black Ratio") + theme(plot.title = element_text(hjust = 0.5))
```

## Get individual hue count data

> The higher the value of hue, the stronger the filter effect.  For an example,

```{r}
## Get the length of data, hue_range
len <- dim(da)[2]

## empty matrix that will get the frequency data
avo_frequency <- as.matrix(0:49)

for (i in colnames(da)){
  if (i == "hue_index"){
   }
  else{
   ##Do something
   temp <- getIncre(as.matrix(da[[i]]))
   temp <- as.matrix(temp)
   ##get the increment
   avo_frequency  <- cbind(avo_frequency ,temp)
   ##start adding them
  }
}

#now add column names
d_avo <- data.frame(avo_frequency)
colnames(d_avo) <- colnames(da)

```

- `d_avo_frequency` (see below) contains frequency of pixles whose color changed when `hue` was
incremented by 1

```{r }
#convert wide to long extract data attribute 
#and create new columns
dff <-(melt(d_avo, id.vars=c("hue_index")))
dff$block<- substring(dff$variable,0,1)
dff$treatment <- substring(dff$variable,2,2)
dff$id <- substring(dff$variable,3,4)
d_avo_frequency <- dff[,c("block","treatment","id","hue_index","value")]
```

- `20` HUE indidate percent of avocado whose that became black when hue was changed 
from 19 to 20. (bad)

- `25` HUE indicate percent of avocado that turn black when hue was increased from
24 to 25. (still good)

- `30` HUE indicate percent of avocado that turn black when hue changed from 29 to 30.

```{r, message=FALSE, warning=FALSE, fig.cap="Hue frequency"}
d_avo_frequency %>% ggplot(aes(x = hue_index, y = value, color=treatment)) + 
  geom_point(aes(color = treatment)) + facet_wrap(~block,ncol = 1) + xlim(20,35) +
  xlab("Hue") + ylab("Count") + theme(plot.title = element_text(hjust = 0.5))
```

## Create avocado pdf

```{r}
#now df_avo cotains percent of pixles whose color changed when 
# head(d_avo)

## empty matrix that will get the frequency data
avo_pdf <- as.matrix(0:49)

for (i in colnames(d_avo)){
  if (i == "hue_index"){
   }
  else{
   ##Do something
   temp <- sum(d_avo[[i]])
   temp <- d_avo[[i]]/temp
   ##start adding them
   avo_pdf   <- cbind(avo_pdf  ,temp)
  }
}

#now add column names
d_avo_pdf <- data.frame(avo_pdf)
colnames(d_avo_pdf) <- colnames(da)
# head(d_avo_pdf)

#convert wide to long extract data attribute 
#and create new columns
dff <-(melt(d_avo_pdf, id.vars=c("hue_index")))
dff$block<- substring(dff$variable,0,1)
dff$treatment <- substring(dff$variable,2,2)
dff$id <- substring(dff$variable,3,4)
d_avo_pdf_long <- dff[,c("block","treatment","id","hue_index","value")]
# head(d_avo_pdf_long)
```

# Create sample data based on the pdf

```{r, message=FALSE, warning=FALSE, fig.cap="Empirical distribution",}
BT <- d_avo_pdf_long %>% filter(block=="B" & treatment =="t") %>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE)) 

BC <- d_avo_pdf_long %>% filter(block=="B" & treatment =="c")%>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE))

NT <- d_avo_pdf_long %>% filter(block=="N" & treatment =="t")%>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE)) 

NC <- d_avo_pdf_long %>% filter(block=="N" & treatment =="c")%>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE)) 

JT <- d_avo_pdf_long %>% filter(block=="J" & treatment =="t")%>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE)) 

JC <- d_avo_pdf_long %>% filter(block=="J" & treatment =="c")%>% 
      group_by(hue_index) %>% dplyr::summarize(Mean = mean(value, na.rm=TRUE)) 

#treatment
s1 <- get_ind_data(BT)
t1 <- data.frame(block="B", control = "treatment", value = s1)

s2 <- get_ind_data(JT)
t2 <- data.frame(block="J", control = "treatment", value = s2)

s3 <- get_ind_data(NT)
t3 <- data.frame(block="N", control = "treatment", value = s3)

three_treats <- rbind(t1,t2,t3)

#control
s4 <- get_ind_data(BC)
t4 <- data.frame(block="B", control = "control", value = s4)

s5 <- get_ind_data(JC)
t5 <- data.frame(block="J", control = "control", value = s5)

s6 <- get_ind_data(NC)
t6 <- data.frame(block="N", control = "control", value = s6)
three_control <- rbind(t4,t5,t6)

data <- rbind(three_treats,three_control)

p1 <- data%>% ggplot(.,aes(x=value)) +
           geom_density(aes(fill=control),adjust=1.5,alpha=0.3)  +
           facet_wrap(~block, ncol = 1) +
           xlim(20, 45) +
           theme(
                  legend.position="top",
                  panel.spacing = unit(0.1, "lines"),
                  axis.ticks.x=element_blank(),
                  plot.title = element_text(hjust = 0.5)
                ) +
        ggtitle("emprical pdf") +
        xlab("Hue") + ylab("Probability") 
    

p2 <- data %>% ggplot(.,aes(x=value, colour = control)) + stat_ecdf() +
  facet_wrap(~block, ncol = 1) +
           xlim(20, 45) +
           theme(
                  legend.position="top",
                  panel.spacing = unit(0.1, "lines"),
                  axis.ticks.x=element_blank(),
                  plot.title = element_text(hjust = 0.5)
                ) +
        ggtitle("empirical cdf") +
        xlab("Hue") + ylab("Probability")

p1 | p2 
```


### Test based on the maximum distance between empirical distributions

```{r, message=FALSE, warning=FALSE, out.height="200%", fig.cap="Result of KS permutation test"}
#control <- getIncre(df1)
#treatment <- getIncre(df2)
control <- BT$Mean
treatment <- BC$Mean

#sharp null distribution
par(mfrow=c(3,1))
invisible(capture.output(get_ks_permutation(BT$Mean,BC$Mean,5000)))
invisible(capture.output(get_ks_permutation(JT$Mean,JC$Mean,5000)))
invisible(capture.output(get_ks_permutation(NT$Mean,NC$Mean,5000)))
par(mfrow=c(1,1))
Z <- c(control,treatment)
n <- length(control )
m <- length(treatment)
N <- length(Z)
```

> No significant treatment effect was observed

### hue_turn 

```{r}
#raw data converted to long format
# head(d_avo_raw) 

#arrange the data by block, treatement, id and hue_index
d <- as.data.table(d_avo_raw)
d <- d %>% filter(hue_index > 18 & hue_index < 44)
db <- d[order(rank(block), treatment,id,hue_index)]
# head(db)
# dim(db)/36
```

> The following shows two samples shown during the presentation. These
figure sugguest that even hue_turn does not capture the treatment effect of our
interest.

```{r, message=FALSE, warning=FALSE, paged.print=FALSE,fig.cap="Comparision of good and bad avocado black cdf" }
d_temp <- as.data.table(d_avo_raw)
# d_temp
d <- d_temp %>% filter(hue_index > 18 & hue_index < 44) %>% 
           filter (block == "B", id == 2 ) 
db <- d[order(rank(block), treatment,id,hue_index)]

d <- d_temp %>% filter(hue_index > 18 & hue_index < 44) %>% 
           filter (block == "J", id == 11 ) 
dj <- d[order(rank(block), treatment,id,hue_index)]

# tail(db)
# db
p1 <- db %>% ggplot(.,aes(x=hue_index, y = value, colour = treatment)) + geom_line() +
  facet_wrap(~block, ncol = 1) +
           xlim(20, 45) +
           theme(
                  legend.position="top",
                  panel.spacing = unit(0.1, "lines"),
                  axis.ticks.x=element_blank(),
                  plot.title = element_text(hjust = 0.5)
                ) +
        ggtitle("Block B treatment") +
        xlab("Hue") + ylab("Probability")

p2 <- dj %>% ggplot(.,aes(x=hue_index, y = value, colour = treatment)) + geom_line() +
  facet_wrap(~block, ncol = 1) +
           xlim(20, 45) +
           theme(
                  legend.position="top",
                  panel.spacing = unit(0.1, "lines"),
                  axis.ticks.x=element_blank(),
                  plot.title = element_text(hjust = 0.5)
                ) +
        ggtitle("Block J treatment ") +
        xlab("Hue") + ylab("Probability")

p1/p2
```


### Code for mannual confirmation

```{r, fig.cap="Example of abrupt change detection" }
d <- as.data.table(d_avo_raw)
d <- d %>% filter(hue_index > 18 & hue_index < 44)
db <- d[order(rank(block), treatment,id,hue_index)]
i = 23
id <- i
start <- id + (id-1)*24
end <- id*25
d_temp <- as.data.frame(db[start:end,])
dg <- d_temp$value

d_temp[1,"block"]
#abrupt change detection point
dg.amoc=cpt.mean(dg)
v = cpts(dg.amoc)
par(mfrow=c(1,1))
plot(dg,xaxt='n', ,
        xlab="Hue",
        ylab="Black Ratio")
abline(v=v,col="red")
```


```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
d <- as.data.table(d_avo_raw)
d <- d %>% filter(hue_index > 18 & hue_index < 44)
db <- d[order(rank(block), treatment,id,hue_index)]

b_result <- data.frame(block = numeric(0),avocado_number= numeric(0),
                       treatment = numeric(0),
                       hue_turn = numeric(0) )



#hue index starts from 19
# par(mfrow=c(3,4))
i = 1
for (i in 1:36){
  start <- i + (i-1)*24
  end <- i*25
  d_temp <- as.data.frame(db[start:end,])
  dg <- d_temp$value
  block <- d_temp[1,"block"]
  treatment <- d_temp[1,"treatment"]
  id <- d_temp[1,"id"]
  #abrupt change detection point
  dg.amoc=cpt.mean(dg)
  v = cpts(dg.amoc)
  # print(v)
  abrupt <- v + 19
  # print(abrupt)
  b_result[i,] <- c(block,id,treatment,abrupt)
  # plot(dg)
  # abline(v=v,col="red")
}
```

# Power Calculation

## Hue turn

```{r}
b_result$hue_turn <- as.numeric(as.character(b_result$hue_turn))
setDT(b_result)
```

```{r}
mean_hue <- b_result[, .(mean_hue_turn = mean(hue_turn)), keyby = .(treatment)]
mean_hue
```

```{r}
sd_hue <- b_result[, .(sd_hue_turn = sd(hue_turn)), keyby = .(treatment)]
sd_hue
```

```{r}
power_test_t <- function(
  mean_control = 24.22222, 
  mean_treat = 25.11111, 
  sd_control = 1.895988, 
  sd_treat = 2.632129,
  number_per_condition = 18, 
  power_loops = 100, 
  ri_loops = 100, 
  verbose = TRUE) { 

    p_values <- NA   
    ri <- NA 
    d <- data.table()
  
    d[ , condition := rep(c('control', 'treatment'), each = number_per_condition)]  
  
    for(power_loop in 1:power_loops) { 
      if(verbose == TRUE) {
        if(power_loop %% 10 == 0) {
          cat(sprintf('Loop Number: %.0f\n', power_loop))
        }
      } 
      
      # d[condition == 'control',   Y := rnorm(.N, mean = mean_control, sd = sd_control)]
      # d[condition == 'treatment', Y := rnorm(.N, mean = mean_treat, sd = sd_treat)]
      # 
      # ate <- d[ , mean(Y), keyby = condition][ , diff(V1)]
      # 
      # for(ri_loop in 1:ri_loops) { 
      #   ri[ri_loop] <- d[ , mean(Y), keyby = sample(condition)][ , diff(V1)]
      #   }
      # 
      # p_values[power_loop] <- mean(abs(ri) > abs(ate))
      
      p_values[power_loop] <- t.test(
        x = rnorm(number_per_condition, mean = mean_control, sd = sd_control), 
        y = rnorm(number_per_condition, mean = mean_treat, sd = sd_treat)
      )$p.value
    }
      
    return(list(
      'p_values' = p_values, 
      'power' = mean(p_values < 0.05)
      ))
}
```


```{r}
power_test_t()$power
```

### Analysis

> For hue_turn, the statistical power is 0.23, which is small because the mean of hue_turn in each group is close to each other.


## Weight

### read a csv file

```{r}
d_weight <- fread('./data/avo_agg.csv')
d_weight
```


### calculate the difference between before and after 
```{r}
d_weight[, tau := Pre_weight - After]
d_weight
```

```{r}
d_weight_mean <- d_weight[, .(mean_weight = mean(tau)), keyby = .(treatment)]
d_weight_mean
```

```{r}
d_weight_sd <- d_weight[, .(sd_weight = sd(tau)), keyby = .(treatment)]
d_weight_sd
```

```{r}
power_test_t <- function(
  mean_control = 17.94444, 
  mean_treat = 7.00000, 
  sd_control = 3.857342, 
  sd_treat = 8.022028,
  number_per_condition = 18, 
  power_loops = 100, 
  ri_loops = 100, 
  verbose = TRUE) { 

    p_values <- NA   
    ri <- NA 
    d <- data.table()
  
    d[ , condition := rep(c('control', 'treatment'), each = number_per_condition)]  
  
    for(power_loop in 1:power_loops) { 
      if(verbose == TRUE) {
        if(power_loop %% 10 == 0) {
          cat(sprintf('Loop Number: %.0f\n', power_loop))
        }
      } 
      
      # d[condition == 'control',   Y := rnorm(.N, mean = mean_control, sd = sd_control)]
      # d[condition == 'treatment', Y := rnorm(.N, mean = mean_treat, sd = sd_treat)]
      # 
      # ate <- d[ , mean(Y), keyby = condition][ , diff(V1)]
      # 
      # for(ri_loop in 1:ri_loops) { 
      #   ri[ri_loop] <- d[ , mean(Y), keyby = sample(condition)][ , diff(V1)]
      #   }
      # 
      # p_values[power_loop] <- mean(abs(ri) > abs(ate))
      
      p_values[power_loop] <- t.test(
        x = rnorm(number_per_condition, mean = mean_control, sd = sd_control), 
        y = rnorm(number_per_condition, mean = mean_treat, sd = sd_treat)
      )$p.value
    }
      
    return(list(
      'p_values' = p_values, 
      'power' = mean(p_values < 0.05)
      ))
}
```


```{r}
power_test_t()$power
```

### Analysis

> For weight, the statistical power is 1, which is very high because difference in mean and standard deviation between the control group and the treatment group is very large.


