---
title: "Power_calc_banana"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(data.table)
library(stargazer)
library(dplyr)
library(tidyverse)
library(patchwork)
library(sandwich)
library(reshape2)
library(lmtest)#coeftest()
library(bcp)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
getChangeDate <- function(df) {
  #this function is called after 
  #the 'black_ratio' in the raw data has been
  #coverted to `y` after taking the individual fixed effect into account
  d_temp <- as.data.frame(df)
  dg <- d_temp$y
  set.seed(1000)
  bcp_x <- bcp(dg, return.mcmc = TRUE)
  #plot(bcp_x)
  
  invisible(capture.output(bcp_sum <- as.data.frame(summary(bcp_x))))
  bcp_sum$id <- 1:length(dg)
  #selecting draft_number with posterior probabltiy greater than 0.2
  sel <- bcp_sum[which(bcp_x$posterior.prob > 0.2), ]
  #by this time all the bananas went bad
  sel <- sel[sel$id <9,]
  sel <- sel[which.max(sel$Probability),]
  #get the id
  time_of_change <- time(dg)[sel$id]
  return(time_of_change[1])
}
```

# Banana

## Data Cleaning

```{r}
#data is given in a long format
d <- fread('banana.csv',header=FALSE, sep=",")
names(d) <- c('block', 'banana_number', 'treatment', 'day', 'humidity', 
              'temperature', 'black_ratio', 'weight')

```


```{r}
#remove the fixed effect
d[,b_mu:=mean(black_ratio), by = .(block,banana_number)]
d[,y := black_ratio - b_mu]

p1 <- d %>% ggplot(aes(x = day, y = black_ratio)) + 
  geom_point(aes(color = treatment)) + ggtitle("Figure B1(a)") +
  xlab("Day") + ylab("Black Ratio") + theme(plot.title = element_text(hjust = 0.5))

p2 <- d %>% ggplot(aes(x = day, y = y)) + 
  geom_point(aes(color = treatment)) + ggtitle("Figure B1(b)") +
  xlab("Day") + ylab("Black Ratio") + theme(plot.title = element_text(hjust = 0.5))

#p1 / p2
```


```{r}
#included the detrended data 
db <- d[,c('y','block', 'banana_number', 'treatment', 'day', 'humidity', 'temperature','black_ratio','weight')]

#save the result of the detrended data
write.csv(db,"db.csv")

## change the format to process that data
db[,b_name := paste0(block,banana_number)]
db <- db[order(rank(b_name), y)]

## create empty data.frame to store d_turn
b_result <- data.frame(block = numeric(0),banana_number= numeric(0),
                       treatment = numeric(0),
                       d_turn = numeric(0) )

```


## Estimate the treatment effect

### Code for mannual confirmation

```{r}
#try 11 and 24 for demonstration
id <- 25
start <- 1 + (id-1)*10
end <- start + 9
d_temp <- as.data.frame(db[start:end,])

dg <- d_temp$y
dg
set.seed(1000)
bcp_x <- bcp(dg, return.mcmc = TRUE)

#plot(bcp_x,main="Figure B2",
#        xlab="days",
#        ylab="Black Ratio")
```

```{r}
invisible(capture.output(bcp_sum <- as.data.frame(summary(bcp_x))))

bcp_sum$id <- 1:length(dg)
#selecting draft_number with posterior probabltiy greater than 0.2
sel <- as.data.frame(bcp_sum[which(bcp_x$posterior.prob > 0.2), ])
sel <- sel[sel$id <9,]
sel <- sel[which.max(sel$Probability),]
#get the id
time_of_change <- time(dg)[sel$id]
time_of_change[1]
```

### Get the data

```{r}
db[,b_name := paste0(block,banana_number)]
db <- db[order(rank(b_name), y)]

b_result <- data.frame(block = numeric(0),banana_number= numeric(0),
                       treatment = numeric(0),
                       d_turn = numeric(0) )
#has 160 rows
for (val in 1:24)
{
 start <- 1 + (val-1)*10
 end <- (10*val)
 #get the subject information
 val_block   <- db[start,block]
 val_subject <- db[start,banana_number]
 val_treat   <- db[start,treatment]
 
 # # print(paste(start,":",end))
 d_temp <- db[start:end,]
 d_turn <- getChangeDate(d_temp)
 b_result[val,] <- c(val_block,val_subject,val_treat,d_turn)
}

#b_result
```

# Analysis


```{r}
#get control and treatment data
b_control <- as.integer(b_result[b_result[,"treatment"] == "control","d_turn"])
b_treatment <- as.integer(b_result[b_result[,"treatment"] == "treatment","d_turn"])

#put them into a vector
Z <- c(b_control,b_treatment)

#get the length of each vectors
n <- length(b_control)
m <- length(b_treatment)
N <- length(Z)

# Number of permutations
K = 10000

# Test statistic
getMean <- function(A,B) abs(mean(A) - mean(B))

# Test statistic for the observed sample
b_mean <- getMean(b_control,b_treatment)

# Vector of test statistics for each permutation
TT <- vector()

# Permutation test
for(i in 1:K){
 #set.seed(i)
 Z.pi <- sample(Z, N, replace = FALSE)
 TT[i] <- getMean(Z.pi[1:n], Z.pi[(n+1):(n+m)])
}

# Visualising the permuted test statistics
par(bg = 'grey')
hist(TT,main="Figure B3",
        xlab="days",
        ylab="Frequency")
abline(v = b_mean , lwd = 2, col = "yellow")
box()

# approximate p-value
b_pvalue <- mean(TT>b_mean )
```


### statistical power analysis for d_turn

```{r}
mean(b_control)
```

```{r}
mean(b_treatment)
```

```{r}
sd(b_control)
```

```{r}
sd(b_treatment)
```

```{r}
p_values <- NA 
ri <- NA 
d <- data.table()

d[ , condition := rep(c('control', 'treatment'), each = 12)]

for(power_loop in 1:100) { 
  if(power_loop %% 10 == 0) cat(sprintf('Loop Number: %.0f\n', power_loop))
  d[condition == 'control', Y := rnorm(.N, mean = mean(b_control), sd = sd(b_control))]
  d[condition == 'treatment', Y := rnorm(.N, mean = mean(b_treatment), sd = sd(b_treatment))]
  
  ate <- d[ , mean(Y), keyby = condition][ , diff(V1)]
  
  for(ri_loop in 1:100) { 
    ri[ri_loop] <- d[ , mean(Y), keyby = sample(condition)][ , diff(V1)]
    }
  
  p_values[power_loop] <- mean(abs(ri) > abs(ate))
  
}

hist(
  x = p_values, 
  col = 'black', 
  main = 'Histogram of p-values', 
  xlab = 'p-values')
```

```{r}
mean(p_values < 0.05)
```

### Analysis

> The statistical power is 0.32, which is not large because our sample size is not large.


### Statistical Power Calculation for weight and black_ratio

#### read the table
```{r}
#data is given in a long format
d <- fread('banana.csv',header=FALSE, sep=",")
names(d) <- c('block', 'banana_number', 'treatment', 'day', 'humidity', 
              'temperature', 'black_ratio', 'weight')

```

```{r}
head(d)
```

## Weight

### create `d_weight`

```{r}
d_weight <- d[day == 1 | day ==10, c('block', 'treatment', 'banana_number', 'day', 'weight')]
d_weight
```

#### create `d_weight_tau`  and calculate the difference of weight between day 10 and day 1

```{r}
d_weight_tau <- d_weight[, .(difference = -diff(weight)), by = .(block, banana_number, treatment)]
d_weight_tau
```

#### calculate mean and sd for control and treatment

```{r}
mean_weight_control <- d_weight_tau[treatment == 'control', mean(difference)]
mean_weight_control
```

```{r}
std_weight_control <- d_weight_tau[treatment == 'control', sd(difference)]
std_weight_control
```


```{r}
mean_weight_treatment <- d_weight_tau[treatment == 'treatment', mean(difference)]
mean_weight_treatment
```

```{r}
std_weight_treatment <- d_weight_tau[treatment == 'treatment', sd(difference)]
std_weight_treatment
```

### power calculation

```{r}
power_test_t <- function(
  mean_control = mean_weight_control, 
  mean_treat = mean_weight_treatment, 
  sd_control = std_weight_control, 
  sd_treat = std_weight_treatment,
  number_per_condition = 12, 
  power_loops = 100, 
  ri_loops = 100, 
  verbose = TRUE) { 

    p_values <- NA   
    ri <- NA 
    d <- data.table()
  
    d[ , condition := rep(c('control', 'treatment'), each = number_per_condition)]  
  
    for(power_loop in 1:power_loops) { 
      if(verbose == TRUE) {
        if(power_loop %% 10 == 0) {
          cat(sprintf('Loop Number: %.0f\n', power_loop))
        }
      } 
      
      # d[condition == 'control',   Y := rnorm(.N, mean = mean_control, sd = sd_control)]
      # d[condition == 'treatment', Y := rnorm(.N, mean = mean_treat, sd = sd_treat)]
      # 
      # ate <- d[ , mean(Y), keyby = condition][ , diff(V1)]
      # 
      # for(ri_loop in 1:ri_loops) { 
      #   ri[ri_loop] <- d[ , mean(Y), keyby = sample(condition)][ , diff(V1)]
      #   }
      # 
      # p_values[power_loop] <- mean(abs(ri) > abs(ate))
      
      p_values[power_loop] <- t.test(
        x = rnorm(number_per_condition, mean = mean_control, sd = sd_control), 
        y = rnorm(number_per_condition, mean = mean_treat, sd = sd_treat)
      )$p.value
    }
      
    return(list(
      'p_values' = p_values, 
      'power' = mean(p_values < 0.05)
      ))
}
```

```{r}
power_test_t()$power
```

### Analysis

> Statistical power for weight is 0.28, which is not large because of our small sample size.


## Black_ratio

### create `d_black_ratio`

```{r}
d_black_ratio <- d[day == 1 | day ==10, c('block', 'treatment', 'banana_number', 'day', 'black_ratio')]
head(d_black_ratio)
```

### create `d_black_ratio_tau`

```{r}
d_black_ratio_tau <- d_black_ratio[, .(difference = diff(black_ratio)), by = .(block, banana_number, treatment)]
d_black_ratio_tau
```

#### calculate mean and standard deviation for control and treatment

```{r}
mean_black_ratio_control <- d_black_ratio_tau[treatment == 'control', mean(difference)]
mean_black_ratio_control
```


```{r}
std_black_ratio_control <- d_black_ratio_tau[treatment == 'control', sd(difference)]
std_black_ratio_control
```

```{r}
mean_black_ratio_treatment <- d_black_ratio_tau[treatment == 'treatment', mean(difference)]
mean_black_ratio_treatment
```

```{r}
std_black_ratio_treatment <- d_black_ratio_tau[treatment == 'treatment', sd(difference)]
std_black_ratio_treatment
```

### calculate statistical power 

```{r}
power_test_t <- function(
  mean_control = mean_black_ratio_control, 
  mean_treat = mean_black_ratio_treatment, 
  sd_control = std_black_ratio_control, 
  sd_treat = std_black_ratio_treatment,
  number_per_condition = 12, 
  power_loops = 100, 
  ri_loops = 100, 
  verbose = TRUE) { 

    p_values <- NA   
    ri <- NA 
    d <- data.table()
  
    d[ , condition := rep(c('control', 'treatment'), each = number_per_condition)]  
  
    for(power_loop in 1:power_loops) { 
      if(verbose == TRUE) {
        if(power_loop %% 10 == 0) {
          cat(sprintf('Loop Number: %.0f\n', power_loop))
        }
      } 
      
      # d[condition == 'control',   Y := rnorm(.N, mean = mean_control, sd = sd_control)]
      # d[condition == 'treatment', Y := rnorm(.N, mean = mean_treat, sd = sd_treat)]
      # 
      # ate <- d[ , mean(Y), keyby = condition][ , diff(V1)]
      # 
      # for(ri_loop in 1:ri_loops) { 
      #   ri[ri_loop] <- d[ , mean(Y), keyby = sample(condition)][ , diff(V1)]
      #   }
      # 
      # p_values[power_loop] <- mean(abs(ri) > abs(ate))
      
      p_values[power_loop] <- t.test(
        x = rnorm(number_per_condition, mean = mean_control, sd = sd_control), 
        y = rnorm(number_per_condition, mean = mean_treat, sd = sd_treat)
      )$p.value
    }
      
    return(list(
      'p_values' = p_values, 
      'power' = mean(p_values < 0.05)
      ))
}
```


```{r}
power_test_t()$power
```

**result**

Statistical power for black_ratio is 0.51, which is not small although our sample size is small.



## Simple regression for weight

```{r}
mod_weight <- d_weight_tau[, lm(difference ~ treatment + as.factor(block))]
mod_weight$vcovHC_ <- vcovHC(mod_weight)
coeftest(mod_weight, vcov. = mod_weight$vcovHC_)
```

### cohen's d

```{r}
library('effsize')
```

```{r}
d_weight_tau[, cohen.d(-difference, treatment)]
```

### Analysis

> From the regression, treatment effect is 1.973g with a p-value of 0.206, which is not statistically significant. When we calculated cohen's d, it's 0.346. Therefore, we concluded that the treatment effect for weight is not pratically significant.




## simple regression for black_ratio

```{r}
mod_black_ratio <- d_black_ratio_tau[, lm(difference ~ treatment + as.factor(block))]
mod_black_ratio$vcovHC_ <- vcovHC(mod_black_ratio)
coeftest(mod_black_ratio, vcov. = mod_black_ratio$vcovHC_)
```

```{r}
d_black_ratio_tau[, cohen.d(difference ~ treatment)]
```

### Analysis

> From the regression treatment estimate is -0.04 with a p-value of 0.031, which is statistically significant. When we calculated cohen's d, it's 0.856. We can say it's practically significant.

